---
title: "Build Population Estimate Data"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook creates the requested population estimates.

## Dependencies
This notebook requires a number of different `R` packages:

```{r load-packages}
# tidyverse packages
library(dplyr)         # data wrangling
library(readr)         # working with csv data
library(stringr)       # string tools

# spatial packages
library(areal)         # interpolation
library(sf)            # working with spatial data
library(tidycensus)    # census api access
library(tigris)        # tiger/line api access

# other packages
library(here)          # file path management
library(testthat)      # unit testing
```

We also use a function for unit testing ID numbers:

```{r load-functions}
source(here("source", "unique_id.R"))
```

## Create Demographic Data, 1940-2000
These decennial census data were obtained from two sources. The tract-level shapefiles were obtained from IPUMS' [NHGIS](https://www.nhgis.org) database. They come for the entire U.S. (or as much of the U.S. as was tracted at that point - full tract coverage is relatively recent). They were merged with tract-level data obtained from [Social Explorer](http://socialexplorer.com) that was already clean and ready to use for each decade.

### 1940
First, we need to load the shapefile geometry:

```{r load-1940-shapefile}
st_read(here("data", "spatial", "STL_DEMOGRAPHICS_tracts40", "STL_DEMOGRAPHICS_tracts40.shp"),
        stringsAsFactors = FALSE) %>%
  st_transform(crs = 26915) -> stl40
```

Next, we need to load the census data and combine it with the spatial data. We need to create the `TRACTID` variable out of a larger variable named `Geo_Name`. A unit test is included to ensure that the `TRACTID` variable we are creating uniquely identifies observations:

```{r add-1940-pop}
read_csv(here("data", "tabular", "STL_DEMOGRAPHICS_pop40.csv")) %>%
  select(Geo_Name, SE_T001_001) %>%
  rename(TRACTID = Geo_Name, 
         pop40 = SE_T001_001) %>%
  mutate(TRACTID = str_pad(string = TRACTID, width = 5, side = "left", pad = "0")) -> pop40

# unit test
pop40 %>% unique_id(TRACTID) -> idUnique
expect_equal(idUnique, TRUE)

# join data
stl40 <- left_join(stl40, pop40, by = "TRACTID")
```

Finally, we'll use a technique called [areal weighted interpolation](https://slu-opengis.github.io/areal/articles/areal-weighted-interpolation.html) to produce estimates at the neighborhood level. We'll import the neighborhood data, re-project it so that it matches the projection used for the 1940 tract boundaries, subset it so that we have only the needed columns and only residential neighborhoods (large parks removed), and then interpolate all of the tract data into neighborhoods.

```{r interpolate-1940}
# interpolate
st_read(here("data", "spatial", "nhood", "BND_Nhd88_cw.shp"), stringsAsFactors = FALSE) %>%
  st_transform(crs = 26915) %>%
  select(NHD_NAME, NHD_NUM) %>%
  filter(NHD_NUM <= 79) %>%
  aw_interpolate(tid = NHD_NUM, source = stl40, sid = TRACTID, 
                 weight = "sum", output = "tibble", 
                 extensive = "pop40") -> nhood40

# unit test
expect_equal(aw_verify(source = stl40, sourceValue = pop40, result = nhood40, resultValue = pop40), TRUE)

# clean-up enviornment
rm(pop40, stl40)
```

For tracts that straddle one of the large parks, their entire population is allocated into the appropriate adjacent neighborhood. We confirm that the entire city's population using a unit test with the `aw_verify()` function. As long as `aw_verify()` returns `TRUE`, we know that each resident has been allocated. We wrap this in a unit test so that the code errors out if this assumption is not met.

### 1950
For the remainder of the decennial census data, I'm going to use the same workflow but condense the code.

The only major change with this decade is that the tract ID numbers already have the appropriate number of leading zeros, so the code to add them can be omitted.

```{r create-1950-estimates}
# read in 1950 era tract boundaries, re-project
st_read(here("data", "spatial", "STL_DEMOGRAPHICS_tracts50", "STL_DEMOGRAPHICS_tracts50.shp"),
        stringsAsFactors = FALSE) %>%
  st_transform(crs = 26915) -> stl50

# read in 1950 census counts, clean
read_csv(here("data", "tabular", "STL_DEMOGRAPHICS_pop50.csv")) %>%
  select(Geo_Name, SE_T001_001) %>%
  rename(TRACTID = Geo_Name, 
         pop50 = SE_T001_001) -> pop50

# unit test
pop50 %>% unique_id(TRACTID) -> idUnique
expect_equal(idUnique, TRUE)

# join data
stl50 <- left_join(stl50, pop50, by = "TRACTID")

# interpolate to neighborhoods
st_read(here("data", "spatial", "nhood", "BND_Nhd88_cw.shp"), stringsAsFactors = FALSE) %>%
  st_transform(crs = 26915) %>%
  select(NHD_NUM) %>%
  filter(NHD_NUM <= 79) %>%
  aw_interpolate(tid = NHD_NUM, source = stl50, sid = TRACTID, 
                 weight = "sum", output = "tibble", 
                 extensive = "pop50") -> nhood50

# unit test
expect_equal(aw_verify(source = stl50, sourceValue = pop50, result = nhood50, resultValue = pop50), TRUE)

# clean-up enviornment
rm(pop50, stl50)
```

### 1960
We'll repeat this process for the 1960s era data.

One issue with the 1960s data is that the `TRACTID` in the 1960 census data needs to be parsed out of a larger variable `Geo_FIPS`. This is addressed in the second pipeline.

```{r create-1960-estimates}
# read in 1960 era tract boundaries, re-project
st_read(here("data", "spatial", "STL_DEMOGRAPHICS_tracts60", "STL_DEMOGRAPHICS_tracts60.shp"),
        stringsAsFactors = FALSE) %>%
  st_transform(crs = 26915) -> stl60

# read in 1960 census counts, clean
read_csv(here("data", "tabular", "STL_DEMOGRAPHICS_pop60.csv")) %>%
  select(Geo_FIPS, SE_T001_001) %>%
  rename(TRACTID = Geo_FIPS, 
         pop60 = SE_T001_001) %>%
  mutate(TRACTID = str_sub(TRACTID, 7, 11)) -> pop60

# unit test
pop60 %>% unique_id(TRACTID) -> idUnique
expect_equal(idUnique, TRUE)

# join data
stl60 <- left_join(stl60, pop60, by = "TRACTID")

# interpolate to neighborhoods
st_read(here("data", "spatial", "nhood", "BND_Nhd88_cw.shp"), stringsAsFactors = FALSE) %>%
  st_transform(crs = 26915) %>%
  select(NHD_NUM) %>%
  filter(NHD_NUM <= 79) %>%
  aw_interpolate(tid = NHD_NUM, source = stl60, sid = TRACTID, 
                 weight = "sum", output = "tibble", 
                 extensive = "pop60") -> nhood60

# unit test
expect_equal(aw_verify(source = stl60, sourceValue = pop60, result = nhood60, resultValue = pop60), TRUE)

# clean-up enviornment
rm(pop60, stl60)
```

### 1970
We'll repeat this process for the 1970s era data.

```{r create-1970-estimates}
# read in 1970 era tract boundaries, re-project
st_read(here("data", "spatial", "STL_DEMOGRAPHICS_tracts70", "STL_DEMOGRAPHICS_tracts70.shp"),
        stringsAsFactors = FALSE) %>%
  st_transform(crs = 26915) -> stl70

# read in 1970 census counts, clean
read_csv(here("data", "tabular", "STL_DEMOGRAPHICS_pop70.csv")) %>%
  select(Geo_TractCode, SE_T001_001) %>%
  rename(TRACTID = Geo_TractCode, 
         pop70 = SE_T001_001) -> pop70

# unit test
pop70 %>% unique_id(TRACTID) -> idUnique
expect_equal(idUnique, TRUE)

# join data
stl70 <- left_join(stl70, pop70, by = "TRACTID")

# interpolate to neighborhoods
st_read(here("data", "spatial", "nhood", "BND_Nhd88_cw.shp"), stringsAsFactors = FALSE) %>%
  st_transform(crs = 26915) %>%
  select(NHD_NUM) %>%
  filter(NHD_NUM <= 79) %>%
  aw_interpolate(tid = NHD_NUM, source = stl70, sid = TRACTID, 
                 weight = "sum", output = "tibble", 
                 extensive = "pop70") -> nhood70

# unit test
expect_equal(aw_verify(source = stl70, sourceValue = pop70, result = nhood70, resultValue = pop70), TRUE)

# clean-up enviornment
rm(pop70, stl70)
```

### 1980
We'll repeat this process for the 1980s era data.

```{r create-1980-estimates}
# read in 1980 era tract boundaries, re-project
st_read(here("data", "spatial", "STL_DEMOGRAPHICS_tracts80", "STL_DEMOGRAPHICS_tracts80.shp"),
        stringsAsFactors = FALSE) %>%
  st_transform(crs = 26915) -> stl80

# read in 1980 census counts, clean
read_csv(here("data", "tabular", "STL_DEMOGRAPHICS_pop80.csv")) %>%
  select(Geo_TRACT6, SE_T001_001) %>%
  rename(TRACTID = Geo_TRACT6, 
         pop80 = SE_T001_001) -> pop80

# unit test
pop80 %>% unique_id(TRACTID) -> idUnique
expect_equal(idUnique, TRUE)

# join data
stl80 <- left_join(stl80, pop80, by = "TRACTID")

# interpolate to neighborhoods
st_read(here("data", "spatial", "nhood", "BND_Nhd88_cw.shp"), stringsAsFactors = FALSE) %>%
  st_transform(crs = 26915) %>%
  select(NHD_NUM) %>%
  filter(NHD_NUM <= 79) %>%
  aw_interpolate(tid = NHD_NUM, source = stl80, sid = TRACTID, 
                 weight = "sum", output = "tibble", 
                 extensive = "pop80") -> nhood80

# unit test
expect_equal(aw_verify(source = stl80, sourceValue = pop80, result = nhood80, resultValue = pop80), TRUE)

# clean-up enviornment
rm(pop80, stl80)
```

### 1990
We'll repeat this process for the 1990s era data.

```{r create-1990-estimates}
# read in 1990 era tract boundaries, re-project
st_read(here("data", "spatial", "STL_DEMOGRAPHICS_tracts90", "STL_DEMOGRAPHICS_tracts90.shp"),
        stringsAsFactors = FALSE) %>%
  st_transform(crs = 26915) -> stl90

# read in 1990 census counts, clean
get_decennial(geography = "tract", variable = "P0010001", year = 1990, state = 29, county = 510) %>%
  select(GEOID, value) %>%
  rename(TRACTID = GEOID, 
         pop90 = value) -> pop90

# unit test
pop90 %>% unique_id(TRACTID) -> idUnique
expect_equal(idUnique, TRUE)

# join data
stl90 <- left_join(stl90, pop90, by = "TRACTID")

# interpolate to neighborhoods
st_read(here("data", "spatial", "nhood", "BND_Nhd88_cw.shp"), stringsAsFactors = FALSE) %>%
  st_transform(crs = 26915) %>%
  select(NHD_NUM) %>%
  filter(NHD_NUM <= 79) %>%
  aw_interpolate(tid = NHD_NUM, source = stl90, sid = TRACTID, 
                 weight = "sum", output = "tibble", 
                 extensive = "pop90") -> nhood90

# unit test
expect_equal(aw_verify(source = stl90, sourceValue = pop90, result = nhood90, resultValue = pop90), TRUE)

# clean-up enviornment
rm(pop90, stl90)
```

### 2000
We'll repeat this process for the 2000s era data.

```{r create-2000-estimates}
# read in 2000 era tract boundaries, re-project
st_read(here("data", "spatial", "STL_DEMOGRAPHICS_tracts00", "STL_DEMOGRAPHICS_tracts00.shp"),
        stringsAsFactors = FALSE) %>%
  st_transform(crs = 26915) -> stl00

# read in 2000 census counts, clean
read_csv(here("data", "tabular", "STL_DEMOGRAPHICS_pop00.csv")) %>%
  select(Geo_TRACT, SE_T001_001) %>%
  rename(TRACTID = Geo_TRACT, 
         pop00 = SE_T001_001) -> pop00

# unit test
pop00 %>% unique_id(TRACTID) -> idUnique
expect_equal(idUnique, TRUE)

# join data
stl00 <- left_join(stl00, pop00, by = "TRACTID")

# interpolate to neighborhoods
st_read(here("data", "spatial", "nhood", "BND_Nhd88_cw.shp"), stringsAsFactors = FALSE) %>%
  st_transform(crs = 26915) %>%
  select(NHD_NUM) %>%
  filter(NHD_NUM <= 79) %>%
  aw_interpolate(tid = NHD_NUM, source = stl00, sid = TRACTID, 
                 weight = "sum", output = "tibble", 
                 extensive = "pop00") -> nhood00

# unit test
expect_equal(aw_verify(source = stl00, sourceValue = pop00, result = nhood00, resultValue = pop00), TRUE)

# clean-up enviornment
rm(pop00, stl00)
```

### Combine 1940s-2000s Data
Next, we'll join all of the neighborhood estimates we've created so far together into a single object:

```{r join-1940-2000}
left_join(nhood40, nhood50, by = "NHD_NUM") %>%
  left_join(., nhood60, by = "NHD_NUM") %>%
  left_join(., nhood70, by = "NHD_NUM") %>%
  left_join(., nhood80, by = "NHD_NUM") %>%
  left_join(., nhood90, by = "NHD_NUM") %>%
  left_join(., nhood00, by = "NHD_NUM") -> nhoodPop_40_00

# clean up enviornment
rm(nhood40, nhood50, nhood60, nhood70, nhood80, nhood90, nhood00, idUnique, unique_id)
```

## Download Demographic Data, 2010-2017
All data for this era are downloaded from the Census Bureau's API via `tidycensus`. 

### 2010
These data are from the decennial census.

```{r download-2010, results='hide'}
get_decennial(geography = "tract", variables = "P001001", state = 29, county = 510, geometry = TRUE) %>% 
  select(GEOID, NAME, value) %>%
  rename(pop10 = value) -> stl10
```

### 2011
These data are from the 2007-2011 5-year American Community Survey estimates.

```{r download-2011}
get_acs(geography = "tract", year = 2011, variables = "B01003_001", state = 29, county = 510) %>%
  select(GEOID, estimate, moe) %>%
  rename(pop11 = estimate,
         pop11_m = moe) -> stl11
```

### 2012
These data are from the 2008-2012 5-year American Community Survey estimates.

```{r download-2012}
get_acs(geography = "tract", year = 2012, variables = "B01003_001", state = 29, county = 510) %>%
  select(GEOID, estimate, moe) %>%
  rename(pop12 = estimate,
         pop12_m = moe) -> stl12
```

### 2013
These data are from the 2009-2013 5-year American Community Survey estimates.

```{r download-2013}
get_acs(geography = "tract", year = 2013, variables = "B01003_001", state = 29, county = 510) %>%
  select(GEOID, estimate, moe) %>%
  rename(pop13 = estimate,
         pop13_m = moe) -> stl13
```

### 2014
These data are from the 2010-2014 5-year American Community Survey estimates.

```{r download-2014}
get_acs(geography = "tract", year = 2014, variables = "B01003_001", state = 29, county = 510) %>%
  select(GEOID, estimate, moe) %>%
  rename(pop14 = estimate,
         pop14_m = moe) -> stl14
```

### 2015
These data are from the 2011-2015 5-year American Community Survey estimates.

```{r download-2015}
get_acs(geography = "tract", year = 2015, variables = "B01003_001", state = 29, county = 510) %>%
  select(GEOID, estimate, moe) %>%
  rename(pop15 = estimate,
         pop15_m = moe) -> stl15
```

### 2016
These data are from the 2012-2016 5-year American Community Survey estimates.

```{r download-2016}
get_acs(geography = "tract", year = 2016, variables = "B01003_001", state = 29, county = 510) %>%
  select(GEOID, estimate, moe) %>%
  rename(pop16 = estimate,
         pop16_m = moe) -> stl16
```

### 2017
These data are from the 2013-2017 5-year American Community Survey estimates.

```{r download-2017}
get_acs(geography = "tract", year = 2017, variables = "B01003_001", state = 29, county = 510) %>%
  select(GEOID, estimate, moe) %>%
  rename(pop17 = estimate,
         pop17_m = moe) -> stl17
```

### Combine Data
We have these data in a number of different tables, so the next step is to join them together by `GEOID`.

```{r combine-data}
left_join(stl10, stl11, by = "GEOID") %>%
  left_join(., stl12, by = "GEOID") %>%
  left_join(., stl13, by = "GEOID") %>%
  left_join(., stl14, by = "GEOID") %>%
  left_join(., stl15, by = "GEOID") %>%
  left_join(., stl16, by = "GEOID") %>%
  left_join(., stl17, by = "GEOID") %>%
  st_transform(crs = 26915) -> tractPop

# clean up enviornment
rm(stl10, stl11, stl12, stl13, stl14, stl15, stl16, stl17)
```

### Interpolate Neighborhood Data
Next, we'll use the same estimation process we used before on the 2010-2017 data:

```{r interpolate}
# read neighborhood data, re-project, and interpolate
st_read(here("data", "spatial", "nhood", "BND_Nhd88_cw.shp"), stringsAsFactors = FALSE) %>%
  st_transform(crs = 26915) %>%
  select(NHD_NUM) %>%
  filter(NHD_NUM <= 79) %>%
  aw_interpolate(tid = NHD_NUM, source = tractPop, sid = GEOID, 
                 weight = "sum", output = "tibble", 
                 extensive = c("pop10", "pop11", "pop11_m", 
                               "pop12", "pop12_m", "pop13", "pop13_m", 
                               "pop14", "pop14_m", "pop15", "pop15_m", 
                               "pop16", "pop16_m", "pop17", "pop17_m")
                 ) -> nhoodPop_10_17

# unit test
expect_equal(aw_verify(source = tractPop, sourceValue = pop10, result = nhoodPop_10_17, resultValue = pop10), TRUE)
expect_equal(aw_verify(source = tractPop, sourceValue = pop11, result = nhoodPop_10_17, resultValue = pop11), TRUE)
expect_equal(aw_verify(source = tractPop, sourceValue = pop12, result = nhoodPop_10_17, resultValue = pop12), TRUE)
expect_equal(aw_verify(source = tractPop, sourceValue = pop13, result = nhoodPop_10_17, resultValue = pop13), TRUE)
expect_equal(aw_verify(source = tractPop, sourceValue = pop14, result = nhoodPop_10_17, resultValue = pop14), TRUE)
expect_equal(aw_verify(source = tractPop, sourceValue = pop15, result = nhoodPop_10_17, resultValue = pop15), TRUE)
expect_equal(aw_verify(source = tractPop, sourceValue = pop16, result = nhoodPop_10_17, resultValue = pop16), TRUE)
expect_equal(aw_verify(source = tractPop, sourceValue = pop17, result = nhoodPop_10_17, resultValue = pop17), TRUE)

# clean up enviornment
rm(tractPop)
```

## Combine Historical and Modern Census Data
Next, we'll combine the two data objects to create a single table of census estimates:

```{r join-all-data}
nhoodPop <- left_join(nhoodPop_40_00, nhoodPop_10_17, by = "NHD_NUM")

# clean up enviornment
rm(nhoodPop_40_00, nhoodPop_10_17)
```

## Export
Finally, we'll write the data to a `.csv` file for future analysis.

```{r}
# write output
write_csv(nhoodPop, here("data", "clean", "STL_PopByNhood.csv"))
```
