---
title: "Data Visualizations"
author: "Christopher Prener, Ph.D., Carter Hanford, B.A."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook generates a number of data visualizations, primarily maps, based on the neighborhood population and race estimates.

## Dependencies
This notebook requires a number of different `R` packages:

```{r load-packages}
# tidyverse packages
library(dplyr)         # data wrangling
library(readr)         # working with csv data
library(ggplot2)       # mapping

# spatial packages
library(sf)            # working with spatial data

# other packages
library(here)          # file path management
library(measurements)  # unit converting
```

## Load Source
Here we will load necessary functions for formatting the neighborhood maps and saving them into our results folder.

```{r}
source(here("source", "cp_sequoiaTheme.R"))
source(here("source", "cp_plotSave.R"))
```


## Load Data
In this section, we'll be using the STL Population and Race data by neighborhod located in the clean folder under data.

```{r data-load}
# total population
pop <- read_csv(here("data", "clean", "STL_PopByNhood.csv"))

# race
race <- read_csv(here("data", "clean", "STL_RaceByNhood.csv"))

# neighborhood boundaries
st_read(here("data", "spatial", "nhood", "BND_Nhd88_cw.shp"), 
                    stringsAsFactors = FALSE) %>%
  st_transform(crs = 26915) %>%
  select(NHD_NUM) -> nhood_sf
```

## Merge STL Pop/Race
To generate the appropriate maps, we need to merge the STL population and race by neighborhood data in the clean folder. The join turns generates spatial data that we can map to the city of St. Louis.

```{r join}
# remove excess column
race <- select(race, -NHD_NAME)

# combine
nhood <- left_join(pop, race, by = "NHD_NUM")
nhood <- left_join(nhood_sf, nhood, by = "NHD_NUM") %>%
  filter(NHD_NUM < 80)
```

Next, for formatting purposes, we need to convert from m^2 to km^2:

```{r convert}
nhood %>%
  mutate(sqmi = unclass(st_area(geometry))) %>%
  mutate(sqmi = conv_unit(sqmi, from = "m2", to = "mi2")) -> nhood
```

## Maps - Neighborhoods by Race
Now that we have our spatial data formatted and ready to go, we can start to map the neighborhoods by year.  Using `ggplot`, we'll generate racial composition maps of St. Louis neighborhoods. Starting in 1940, we will generate 15 separate maps, 1940-2010 (decennial), and 2011-2017.  This section will be lengthy and will use recycled code from the first block, so we'll refrain from narrative.

### 1940
First, we'll map the population density of neighborhoods in 1940:

```{r popden-1940}
nhood <- mutate(nhood, den = pop40/sqmi)
min <- round(min(nhood$den), 0)
max <- round(max(nhood$den), 0)
breaks <- c(seq(from = min, to = max, length.out = 5))
breaks[1] <- breaks[1]+1
breaks[5] <- breaks[5]-1

plot_den <- ggplot() +
  geom_sf(data = nhood_sf, fill = "#006a14", color = NA) +
  geom_sf(data = nhood, mapping = aes(fill = den), color = NA) +
  geom_sf(data = nhood_sf, fill = NA, color = "black") +
  scale_fill_distiller(palette = "OrRd", name = "Population\nper Sq. Mi.", trans = "reverse",
                       breaks = breaks) +
  labs(
    title = "Population Density, 1940",
    subtitle = "City of St. Louis Modern Neighborhoods",
    caption = "Map by Christopher Prener, PhD. and Carter Hanford, B.A.\nData via NHGIS and the City of St. Louis") +
  cp_sequoiaTheme(base_size = 20, background = "white", map = TRUE)
```

Next, we'll map the percentage of each neighborhood's population that was not white. A quick note, census data from 1940 does not have a `black` percentage, so we will use `non-white`.

```{r race-1940}
nhood <- mutate(nhood, pct = (nonwhite40/pop40)*100)
min <- round(min(nhood$pct), 0)
max <- round(max(nhood$pct), 0)
breaks <- c(seq(from = min, to = max, length.out = 5))
breaks[1] <- breaks[1]+.1
breaks[5] <- breaks[5]-.1

plot_race <- ggplot() +
  geom_sf(data = nhood_sf, fill = "#006a14", color = NA) +
  geom_sf(data = nhood, mapping = aes(fill = pct), color = NA) +
  geom_sf(data = nhood_sf, fill = NA, color = "black") +
  scale_fill_distiller(palette = "RdPu", name = "% Non-white", trans = "reverse",
                       breaks = breaks) +
  labs(
    title = "Percent Nonwhite, 1940",
    subtitle = "City of St. Louis Modern Neighborhoods",
    caption = "Map by Christopher Prener, PhD. and Carter Hanford, B.A.\nData via NHGIS and the City of St. Louis") +
  cp_sequoiaTheme(base_size = 20, background = "white", map = TRUE)
```

Finally, we'll save both maps:

```{r save-1940}
cp_plotSave(here("results", "Map_PctBlack", "map_1940_nonwhite.png"), plot_race, preset = "lg", dpi = 500)
cp_plotSave(here("results", "Map_PopDensity", "map_1940_pop_density.png"), plot_den, preset = "lg", dpi = 500)
```

### 1950
We'll repeat the same code for 1950:

```{r map-1950}
# population density
nhood <- mutate(nhood, den = pop50/sqmi)
min <- round(min(nhood$den), 0)
max <- round(max(nhood$den), 0)
breaks <- c(seq(from = min, to = max, length.out = 5))
breaks[1] <- breaks[1]+1
breaks[5] <- breaks[5]-1

plot_den <- ggplot() +
  geom_sf(data = nhood_sf, fill = "#006a14", color = NA) +
  geom_sf(data = nhood, mapping = aes(fill = den), color = NA) +
  geom_sf(data = nhood_sf, fill = NA, color = "black") +
  scale_fill_distiller(palette = "OrRd", name = "Population\nper Sq. Mi.", trans = "reverse",
                       breaks = breaks) +
  labs(
    title = "Population Density, 1950",
    subtitle = "City of St. Louis Modern Neighborhoods",
    caption = "Map by Christopher Prener, PhD. and Carter Hanford, B.A.\nData via NHGIS and the City of St. Louis") +
  cp_sequoiaTheme(base_size = 20, background = "white", map = TRUE)

# race
nhood <- mutate(nhood, pct = (black50/pop50)*100)
min <- round(min(nhood$pct), 0)
max <- round(max(nhood$pct), 0)
breaks <- c(seq(from = min, to = max, length.out = 5))
breaks[1] <- breaks[1]+.1
breaks[5] <- breaks[5]-.1

plot_race <- ggplot() +
  geom_sf(data = nhood_sf, fill = "#006a14", color = NA) +
  geom_sf(data = nhood, mapping = aes(fill = pct), color = NA) +
  geom_sf(data = nhood_sf, fill = NA, color = "black") +
  scale_fill_distiller(palette = "RdPu", name = "% African American", trans = "reverse",
                       breaks = breaks) +
  labs(
    title = "Percent Black, 1950",
    subtitle = "City of St. Louis Modern Neighborhoods",
    caption = "Map by Christopher Prener, PhD. and Carter Hanford, B.A.\nData via NHGIS and the City of St. Louis") +
  cp_sequoiaTheme(base_size = 20, background = "white", map = TRUE)

# save
cp_plotSave(here("results", "Map_PctBlack", "map_1950_black.png"), plot_race, preset = "lg", dpi = 500)
cp_plotSave(here("results", "Map_PopDensity", "map_1950_pop_density.png"), plot_den, preset = "lg", dpi = 500)
```

### 1960
We'll repeat the same code for 1960:

```{r map-1960}
# population density
nhood <- mutate(nhood, den = pop60/sqmi)
min <- round(min(nhood$den), 0)
max <- round(max(nhood$den), 0)
breaks <- c(seq(from = min, to = max, length.out = 5))
breaks[1] <- breaks[1]+1
breaks[5] <- breaks[5]-1

plot_den <- ggplot() +
  geom_sf(data = nhood_sf, fill = "#006a14", color = NA) +
  geom_sf(data = nhood, mapping = aes(fill = den), color = NA) +
  geom_sf(data = nhood_sf, fill = NA, color = "black") +
  scale_fill_distiller(palette = "OrRd", name = "Population\nper Sq. Mi.", trans = "reverse",
                       breaks = breaks) +
  labs(
    title = "Population Density, 1960",
    subtitle = "City of St. Louis Modern Neighborhoods",
    caption = "Map by Christopher Prener, PhD. and Carter Hanford, B.A.\nData via NHGIS and the City of St. Louis") +
  cp_sequoiaTheme(base_size = 20, background = "white", map = TRUE)

# race
nhood <- mutate(nhood, pct = (black60/pop60)*100)
min <- round(min(nhood$pct), 0)
max <- round(max(nhood$pct), 0)
breaks <- c(seq(from = min, to = max, length.out = 5))
breaks[1] <- breaks[1]+.1
breaks[5] <- breaks[5]-.3

plot_race <- ggplot() +
  geom_sf(data = nhood_sf, fill = "#006a14", color = NA) +
  geom_sf(data = nhood, mapping = aes(fill = pct), color = NA) +
  geom_sf(data = nhood_sf, fill = NA, color = "black") +
  scale_fill_distiller(palette = "RdPu", name = "% African American", trans = "reverse",
                       breaks = breaks) +
  labs(
    title = "Percent Black, 1960",
    subtitle = "City of St. Louis Modern Neighborhoods",
    caption = "Map by Christopher Prener, PhD. and Carter Hanford, B.A.\nData via NHGIS and the City of St. Louis") +
  cp_sequoiaTheme(base_size = 20, background = "white", map = TRUE)

# save
cp_plotSave(here("results", "Map_PctBlack", "map_1960_black.png"), plot_race, preset = "lg", dpi = 500)
cp_plotSave(here("results", "Map_PopDensity", "map_1960_pop_density.png"), plot_den, preset = "lg", dpi = 500)
```

### 1970
We'll repeat the same code for 1970:

```{r map-1970}
# population density
nhood <- mutate(nhood, den = pop70/sqmi)
min <- round(min(nhood$den), 0)
max <- round(max(nhood$den), 0)
breaks <- c(seq(from = min, to = max, length.out = 5))
breaks[1] <- breaks[1]+1
breaks[5] <- breaks[5]-1

plot_den <- ggplot() +
  geom_sf(data = nhood_sf, fill = "#006a14", color = NA) +
  geom_sf(data = nhood, mapping = aes(fill = den), color = NA) +
  geom_sf(data = nhood_sf, fill = NA, color = "black") +
  scale_fill_distiller(palette = "OrRd", name = "Population\nper Sq. Mi.", trans = "reverse",
                       breaks = breaks) +
  labs(
    title = "Population Density, 1970",
    subtitle = "City of St. Louis Modern Neighborhoods",
    caption = "Map by Christopher Prener, PhD. and Carter Hanford, B.A.\nData via NHGIS and the City of St. Louis") +
  cp_sequoiaTheme(base_size = 20, background = "white", map = TRUE)

# race
nhood <- mutate(nhood, pct = (black70/pop70)*100)
min <- round(min(nhood$pct), 0)
max <- round(max(nhood$pct), 0)
breaks <- c(seq(from = min, to = max, length.out = 5))
breaks[1] <- breaks[1]+.1
breaks[5] <- breaks[5]-.1

plot_race <- ggplot() +
  geom_sf(data = nhood_sf, fill = "#006a14", color = NA) +
  geom_sf(data = nhood, mapping = aes(fill = pct), color = NA) +
  geom_sf(data = nhood_sf, fill = NA, color = "black") +
  scale_fill_distiller(palette = "RdPu", name = "% African American", trans = "reverse",
                       breaks = breaks) +
  labs(
    title = "Percent Black, 1970",
    subtitle = "City of St. Louis Modern Neighborhoods",
    caption = "Map by Christopher Prener, PhD. and Carter Hanford, B.A.\nData via NHGIS and the City of St. Louis") +
  cp_sequoiaTheme(base_size = 20, background = "white", map = TRUE)

# save
cp_plotSave(here("results", "Map_PctBlack", "map_1970_black.png"), plot_race, preset = "lg", dpi = 500)
cp_plotSave(here("results", "Map_PopDensity", "map_1970_pop_density.png"), plot_den, preset = "lg", dpi = 500)
```

### 1980
We'll repeat the same code for 1980:

```{r map-1980}
# population density
nhood <- mutate(nhood, den = pop80/sqmi)
min <- round(min(nhood$den), 0)
max <- round(max(nhood$den), 0)
breaks <- c(seq(from = min, to = max, length.out = 5))
breaks[1] <- breaks[1]+1
breaks[5] <- breaks[5]-1

plot_den <- ggplot() +
  geom_sf(data = nhood_sf, fill = "#006a14", color = NA) +
  geom_sf(data = nhood, mapping = aes(fill = den), color = NA) +
  geom_sf(data = nhood_sf, fill = NA, color = "black") +
  scale_fill_distiller(palette = "OrRd", name = "Population\nper Sq. Mi.", trans = "reverse",
                       breaks = breaks) +
  labs(
    title = "Population Density, 1980",
    subtitle = "City of St. Louis Modern Neighborhoods",
    caption = "Map by Christopher Prener, PhD. and Carter Hanford, B.A.\nData via U.S. Census Bureau and the City of St. Louis") +
  cp_sequoiaTheme(base_size = 20, background = "white", map = TRUE)

# race
nhood <- mutate(nhood, pct = (black80/pop80)*100)
min <- round(min(nhood$pct), 0)
max <- round(max(nhood$pct), 0)
breaks <- c(seq(from = min, to = max, length.out = 5))
breaks[1] <- breaks[1]+.1
breaks[5] <- breaks[5]-.1

plot_race <- ggplot() +
  geom_sf(data = nhood_sf, fill = "#006a14", color = NA) +
  geom_sf(data = nhood, mapping = aes(fill = pct), color = NA) +
  geom_sf(data = nhood_sf, fill = NA, color = "black") +
  scale_fill_distiller(palette = "RdPu", name = "% African American", trans = "reverse",
                       breaks = breaks) +
  labs(
    title = "Percent Black, 1980",
    subtitle = "City of St. Louis Modern Neighborhoods",
    caption = "Map by Christopher Prener, PhD. and Carter Hanford, B.A.\nData via U.S. Census Bureau and the City of St. Louis") +
  cp_sequoiaTheme(base_size = 20, background = "white", map = TRUE)

# save
cp_plotSave(here("results", "Map_PctBlack", "map_1980_black.png"), plot_race, preset = "lg", dpi = 500)
cp_plotSave(here("results", "Map_PopDensity", "map_1980_pop_density.png"), plot_den, preset = "lg", dpi = 500)
```

### 1990
We'll repeat the same code for 1990:

```{r map-1990}
# population density
nhood <- mutate(nhood, den = pop90/sqmi)
min <- round(min(nhood$den), 0)
max <- round(max(nhood$den), 0)
breaks <- c(seq(from = min, to = max, length.out = 5))
breaks[1] <- breaks[1]+1
breaks[5] <- breaks[5]-1

plot_den <- ggplot() +
  geom_sf(data = nhood_sf, fill = "#006a14", color = NA) +
  geom_sf(data = nhood, mapping = aes(fill = den), color = NA) +
  geom_sf(data = nhood_sf, fill = NA, color = "black") +
  scale_fill_distiller(palette = "OrRd", name = "Population\nper Sq. Mi.", trans = "reverse",
                       breaks = breaks) +
  labs(
    title = "Population Density, 1990",
    subtitle = "City of St. Louis Modern Neighborhoods",
    caption = "Map by Christopher Prener, PhD. and Carter Hanford, B.A.\nData via U.S. Census Bureau and the City of St. Louis") +
  cp_sequoiaTheme(base_size = 20, background = "white", map = TRUE)

# race
nhood <- mutate(nhood, pct = (black90/pop90)*100)
min <- round(min(nhood$pct), 0)
max <- round(max(nhood$pct), 0)
breaks <- c(seq(from = min, to = max, length.out = 5))
breaks[1] <- breaks[1]+.2
breaks[5] <- breaks[5]-.1

plot_race <- ggplot() +
  geom_sf(data = nhood_sf, fill = "#006a14", color = NA) +
  geom_sf(data = nhood, mapping = aes(fill = pct), color = NA) +
  geom_sf(data = nhood_sf, fill = NA, color = "black") +
  scale_fill_distiller(palette = "RdPu", name = "% African American", trans = "reverse",
                       breaks = breaks) +
  labs(
    title = "Percent Black, 1990",
    subtitle = "City of St. Louis Modern Neighborhoods",
    caption = "Map by Christopher Prener, PhD. and Carter Hanford, B.A.\nData via U.S. Census Bureau and the City of St. Louis") +
  cp_sequoiaTheme(base_size = 20, background = "white", map = TRUE)

# save
cp_plotSave(here("results", "Map_PctBlack", "map_1990_black.png"), plot_race, preset = "lg", dpi = 500)
cp_plotSave(here("results", "Map_PopDensity", "map_1990_pop_density.png"), plot_den, preset = "lg", dpi = 500)
```

### 2000
We'll repeat the same code for 2000:

```{r map-2000}
# population density
nhood <- mutate(nhood, den = pop00/sqmi)
min <- round(min(nhood$den), 0)
max <- round(max(nhood$den), 0)
breaks <- c(seq(from = min, to = max, length.out = 5))
breaks[1] <- breaks[1]+1
breaks[5] <- breaks[5]-1

plot_den <- ggplot() +
  geom_sf(data = nhood_sf, fill = "#006a14", color = NA) +
  geom_sf(data = nhood, mapping = aes(fill = den), color = NA) +
  geom_sf(data = nhood_sf, fill = NA, color = "black") +
  scale_fill_distiller(palette = "OrRd", name = "Population\nper Sq. Mi.", trans = "reverse",
                       breaks = breaks) +
  labs(
    title = "Population Density, 2000",
    subtitle = "City of St. Louis Modern Neighborhoods",
    caption = "Map by Christopher Prener, PhD. and Carter Hanford, B.A.\nData via U.S. Census Bureau and the City of St. Louis") +
  cp_sequoiaTheme(base_size = 20, background = "white", map = TRUE)

# race
nhood <- mutate(nhood, pct = (black00/pop00)*100)
min <- round(min(nhood$pct), 0)
max <- round(max(nhood$pct), 0)
breaks <- c(seq(from = min, to = max, length.out = 5))
breaks[1] <- breaks[1]+.4
breaks[5] <- breaks[5]-.1

plot_race <- ggplot() +
  geom_sf(data = nhood_sf, fill = "#006a14", color = NA) +
  geom_sf(data = nhood, mapping = aes(fill = pct), color = NA) +
  geom_sf(data = nhood_sf, fill = NA, color = "black") +
  scale_fill_distiller(palette = "RdPu", name = "% African American", trans = "reverse",
                       breaks = breaks) +
  labs(
    title = "Percent Black, 2000",
    subtitle = "City of St. Louis Modern Neighborhoods",
    caption = "Map by Christopher Prener, PhD. and Carter Hanford, B.A.\nData via U.S. Census Bureau and the City of St. Louis") +
  cp_sequoiaTheme(base_size = 20, background = "white", map = TRUE)

# save
cp_plotSave(here("results", "Map_PctBlack", "map_2000_black.png"), plot_race, preset = "lg", dpi = 500)
cp_plotSave(here("results", "Map_PopDensity", "map_2000_pop_density.png"), plot_den, preset = "lg", dpi = 500)
```

### 2010
We'll repeat the same code for 2010:

```{r map-2010}
# population density
nhood <- mutate(nhood, den = pop10/sqmi)
min <- round(min(nhood$den), 0)
max <- round(max(nhood$den), 0)
breaks <- c(seq(from = min, to = max, length.out = 5))
breaks[1] <- breaks[1]+1
breaks[5] <- breaks[5]-1

plot_den <- ggplot() +
  geom_sf(data = nhood_sf, fill = "#006a14", color = NA) +
  geom_sf(data = nhood, mapping = aes(fill = den), color = NA) +
  geom_sf(data = nhood_sf, fill = NA, color = "black") +
  scale_fill_distiller(palette = "OrRd", name = "Population\nper Sq. Mi.", trans = "reverse",
                       breaks = breaks) +
  labs(
    title = "Population Density, 2010",
    subtitle = "City of St. Louis Modern Neighborhoods",
    caption = "Map by Christopher Prener, PhD. and Carter Hanford, B.A.\nData via U.S. Census Bureau and the City of St. Louis") +
  cp_sequoiaTheme(base_size = 20, background = "white", map = TRUE)

# race
nhood <- mutate(nhood, pct = (black10/pop10)*100)
min <- round(min(nhood$pct), 0)
max <- round(max(nhood$pct), 0)
breaks <- c(seq(from = min, to = max, length.out = 5))
breaks[1] <- breaks[1]+.5
breaks[5] <- breaks[5]-.1

plot_race <- ggplot() +
  geom_sf(data = nhood_sf, fill = "#006a14", color = NA) +
  geom_sf(data = nhood, mapping = aes(fill = pct), color = NA) +
  geom_sf(data = nhood_sf, fill = NA, color = "black") +
  scale_fill_distiller(palette = "RdPu", name = "% African American", trans = "reverse",
                       breaks = breaks) +
  labs(
    title = "Percent Black, 2010",
    subtitle = "City of St. Louis Modern Neighborhoods",
    caption = "Map by Christopher Prener, PhD. and Carter Hanford, B.A.\nData via U.S. Census Bureau and the City of St. Louis") +
  cp_sequoiaTheme(base_size = 20, background = "white", map = TRUE)

# save
cp_plotSave(here("results", "Map_PctBlack", "map_2010_black.png"), plot_race, preset = "lg", dpi = 500)
cp_plotSave(here("results", "Map_PopDensity", "map_2010_pop_density.png"), plot_den, preset = "lg", dpi = 500)
```

### 2017
We'll repeat the same code for 2017:

```{r map-2017}
# population density
nhood <- mutate(nhood, den = pop17/sqmi)
min <- round(min(nhood$den), 0)
max <- round(max(nhood$den), 0)
breaks <- c(seq(from = min, to = max, length.out = 5))
breaks[1] <- breaks[1]+1
breaks[5] <- breaks[5]-1

plot_den <- ggplot() +
  geom_sf(data = nhood_sf, fill = "#006a14", color = NA) +
  geom_sf(data = nhood, mapping = aes(fill = den), color = NA) +
  geom_sf(data = nhood_sf, fill = NA, color = "black") +
  scale_fill_distiller(palette = "OrRd", name = "Population\nper Sq. Mi.", trans = "reverse",
                       breaks = breaks) +
  labs(
    title = "Population Density, 2017",
    subtitle = "City of St. Louis Modern Neighborhoods",
    caption = "Map by Christopher Prener, PhD. and Carter Hanford, B.A.\nData via U.S. Census Bureau and the City of St. Louis") +
  cp_sequoiaTheme(base_size = 20, background = "white", map = TRUE)

# race
nhood <- mutate(nhood, pct = (black17/pop17)*100)
min <- round(min(nhood$pct), 0)
max <- round(max(nhood$pct), 0)
breaks <- c(seq(from = min, to = max, length.out = 5))
breaks[1] <- breaks[1]+.3
breaks[5] <- breaks[5]-.4

plot_race <- ggplot() +
  geom_sf(data = nhood_sf, fill = "#006a14", color = NA) +
  geom_sf(data = nhood, mapping = aes(fill = pct), color = NA) +
  geom_sf(data = nhood_sf, fill = NA, color = "black") +
  scale_fill_distiller(palette = "RdPu", name = "% African American", trans = "reverse",
                       breaks = breaks) +
  labs(
    title = "Percent Black, 2017",
    subtitle = "City of St. Louis Modern Neighborhoods",
    caption = "Map by Christopher Prener, PhD. and Carter Hanford, B.A.\nData via U.S. Census Bureau and the City of St. Louis") +
  cp_sequoiaTheme(base_size = 20, background = "white", map = TRUE)

# save
cp_plotSave(here("results", "Map_PctBlack", "map_2017_black.png"), plot_race, preset = "lg", dpi = 500)
cp_plotSave(here("results", "Map_PopDensity", "map_2017_pop_density.png"), plot_den, preset = "lg", dpi = 500)
```
